---
title: "Transit Costs Project"
output: html_notebook
---

The data this week comes from [Transit Costs Project](%5Bhttps://transitcosts.com/).](<https://transitcosts.com/>).)

> Why do transit-infrastructure projects in New York cost 20 times more on a per kilometer basis than in Seoul? We investigate this question across hundreds of transit projects from around the world. We have created a database that spans more than 50 countries and totals more than 11,000 km of urban rail built since the late 1990s. We will also examine this question in greater detail by carrying out six in-depth case studies that take a closer look at unique considerations and variables that aren't easily quantified, like project management, governance, and site conditions.

> The goal of this work is to figure out how to deliver more high-capacity transit projects for a fraction of the cost in countries like the United States. Additionally, we hope that our site will be a useful resource for elected officials, planners, researchers, journalists, advocates, and others interested in contextualizing transit-infrastructure costs and fighting for better projects.

### Data Dictionary

# `transit_cost.csv`

+------------------+-----------------+---------------------------------------------------------------------------------------------+
| variable         | class           | description                                                                                 |
+:=================+:================+:============================================================================================+
| e                | double          | ID                                                                                          |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| country          | character       | Country Code - can be joined against `countrycode` via `ecb` or `iso2c`                     |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| city             | character       | City where transit tunnel is being created                                                  |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| line             | character       | Line name or path                                                                           |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| start_year       | character       | Year started                                                                                |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| end_year         | character       | Year ended (predicted or actual)                                                            |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| rr               | double          | I think this is Railroad (0 or 1), where 1 == Railroad?                                     |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| length           | double          | Length of proposed line in km                                                               |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| tunnel_per       | character       | Percent of line length completed                                                            |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| tunnel           | double          | Tunnel length of line completed in km (can take this divided by length to get `tunnel_per`) |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| stations         | double          | Number of stations where passengers can board/leave                                         |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| source1          | character       | Where was data sourced                                                                      |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| cost             | double          | Cost in millions of local currency                                                          |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| currency         | character       | Currency type                                                                               |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| year             | double          | Midpoint year of construction                                                               |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| ppp_rate         | double          | purchasing power parity (PPP), based on the midpoint of construction                        |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| real_cost        | character       | Real cost in Millions of USD                                                                |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| cost_km_millions | double          | Cost/km in millions of USD                                                                  |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| source2          | character       | Where was data sourced for cost                                                             |
+------------------+-----------------+---------------------------------------------------------------------------------------------+
| reference        | character       | Reference URL for source                                                                    |
+------------------+-----------------+---------------------------------------------------------------------------------------------+

### Code 1

<https://github.com/MartinPons/tidytuesday_MartinPons/blob/master/2021-2/transit_cost.R>

```{r}
library(tidyverse, quietly=TRUE, warn.conflicts=FALSE)
library(glue, quietly=TRUE, warn.conflicts=FALSE)
library(Cairo, quietly=TRUE, warn.conflicts=FALSE)
library(scales, quietly=TRUE, warn.conflicts=FALSE)
library(ggtext, quietly=TRUE, warn.conflicts=FALSE)
```

```{r}
# data
transit_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-05/transit_cost.csv')
transit_cost
```

```{r}
transit_cost <- transit_cost %>% 
  filter(!is.na(e))

# parse numeric variables interpreted as character
transit_cost <- transit_cost %>% 
  mutate(tunnel_per = replace_na(parse_number(tunnel_per), 0),
         real_cost = parse_number(real_cost),
         across(start_year:end_year, as.numeric)) %>% 
  
  # remove outliers_length
  filter(length <= 80)

# fit linear regression cost ~ length
reg <- lm(real_cost ~ length, data = transit_cost)

transit_cost <- 
  transit_cost %>% 
  mutate(fit = predict(reg, transit_cost), 
         error = real_cost - fit)

# dataset for highlighted projects
outliers_error <- 
  transit_cost %>% 
  filter(error < quantile(error, 0.005) | error > quantile(error, 0.995)) %>%
  mutate(years_to_complete = replace_na(end_year - start_year, "Unspecified")) %>% 
  select(e, country, city, line, start_year, years_to_complete, tunnel, real_cost, length, fit)

# addition of formatted text for tags
outliers_error <- outliers_error %>% 
  mutate(text = glue('Project: {line}\nCity: {city} ({country})\nTotal cost: {comma(real_cost, prefix = "$", suffix =  "M")}\nStart: {start_year}\nYears to complete: {years_to_complete}\nDeviation from predicted: {comma(real_cost - fit, prefix = "$", suffix =  "M")}'), 
         hjust = c("right", "right", "right", "right", "left", "left")) 


# dataframe for segements accompaining text for highlighted projects
segments_df <- tribble(
  ~x, ~xend, ~y_shift,
  30, 30, -4000,
  20, 20, 0,
  50, 50, 0,
  72, 72, 4000,
  67, 67, -2600,
  60, 60, -1000
) %>% 
  mutate(e = outliers_error$e)

## creation of path lines connecting hightligthed points to tag text ##

# aesthetics needed for geom_path
outliers_error <- outliers_error %>% 
  left_join(segments_df, by = "e") %>% 
  mutate(y = real_cost - 2300 + y_shift, 
         yend = real_cost + 2300 + y_shift, 
         y2 = real_cost - 300 + y_shift, 
         yend2 = real_cost + 300 + y_shift, 
         x2 = ifelse(hjust == "right", x + 0.3, x - 0.3), 
         x_beggin = ifelse(hjust == "right", length - 0.7, length + 0.7))

# data.frame in format suitable for geom_path
paths_df <- outliers_error %>% 
  mutate(x_middle1 = (length + x) / 2, 
         x_middle2 = (length + x) / 2, 
         y_middle1 = real_cost, 
         y_middle2 = (y + yend) / 2,
         y_end = (y + yend) / 2
         ) %>% 
  select(e, x_beggin, x_middle1, x_middle2, x2, real_cost, y_middle1, y_middle2, y_end) %>% 
  rename(x_end = x2, y_beggin = real_cost) %>% 
  pivot_longer(names_to = c(".value", "point"), names_sep = "_", cols = 2:ncol(.))
```

```{r fig.width=16, fig.height=10}
gg <- transit_cost %>% 
  ggplot(aes(length, real_cost)) + 
  geom_point(color = "grey65", alpha = 0.6) + 
  
  # segments connecting fitted line to highlited points
  geom_segment(data = outliers_error, aes(x = length, xend = length,
                                    y= fit, yend = real_cost - 350),
               color = "grey25") +
  
  # highglighted points
  geom_point(data = outliers_error, color = "#059fff", size = 5, shape = 1, stroke = 1) +
  geom_point(data = outliers_error, color = "white", size = 2.2) +
  
  
  # text and dashed segments indicating the increasing in cost ofr every 10 Km
  annotate(geom = "segment", x = 60, xend = 60,
           y = predict(reg, data.frame(length = 60)),
           yend = predict(reg, data.frame(length = 70)), 
           color = "#bbd1f0", lty = "dashed") +
  
  
  annotate(geom = "segment", x = 60, xend = 70,
           y = predict(reg, data.frame(length = 70)),
           yend = predict(reg, data.frame(length = 70)), 
           color = "#bbd1f0", lty = "dashed") +
  
  annotate(geom = "text", x = 65, y = predict(reg, data.frame(length = 70)) + 900,
           label = glue("Every 10 Km of road increases \n the cost in {comma(reg$coefficients[2], , prefix = '$', suffix =  'M')} on average"),
           color = "#bbd1f0", 
           size = 2.9) +

  # fitted line
  geom_smooth(method = "lm", se = F, color = "#385ee8") +
  
  # blue vertical segments next to tag text
  geom_segment(data = outliers_error, aes(x = x , xend = xend, y = y, yend = yend), color = "#059fff") +
  geom_segment(data = outliers_error, aes(x = x2, 
                                          xend = x2,
                                          y = y2, 
                                          yend = yend2), color = "#059fff") +
  
  # lines connecting highlighted points to vertigal segments
  geom_path(data = paths_df, 
            aes(x, y, group = e), color = "white", 
            linejoin = "bevel", 
            linemitre = 1) +
  
  # tagged text
  geom_text(data = outliers_error, 
            aes(ifelse(hjust == "right",x - 0.5, x + 0.5), (y + yend) / 2, label = text, hjust = hjust), 
            color = "white", 
            size = 3) +

  labs(x = "Length of the line (Km)", 
       y = "Real cost of the project (millions of $)", 
       title = "THE MOST AND LEAST COSTLY TRANSIT-INFRASTRUCTURE PROJECTS AROUND THE WORLD", 
       subtitle = "Each dot represents a project. Highligthed projects are below the 0.5 percentile or above the 99.5 percentile of the  <span style='color:#385ee8'>predicted cost</span>", 
       caption = "Projects are limitied to 80 Km of road or less\nData comes from the Transit Costs Project. Visualization by Martín Pons | @MartinPonsM") +
  
  scale_x_continuous(breaks = seq(0, 80, by = 10)) +
  scale_y_continuous(labels = comma) +

  theme(
    text = element_text(family = "Candara", color = "#9dc6e0"),
    plot.background = element_rect(fill = "grey15"),
    panel.background = element_rect(fill = "grey15"), 
    panel.grid = element_blank(), 
    axis.text = element_text(size = 20, color = "#9dc6e0"), 
    axis.title = element_text(size = 20, color = "#9dc6e0"),
    plot.title = element_text(color = "#ced8f2", size = 21),
    plot.subtitle = element_markdown(color = "#9bb0c9", size = 18),
    plot.caption = element_text(color = "#9bb0c9", size = 16)
) 

gg
```

### Code 2

<https://github.com/ancazugo/TidyTuesday/tree/main/2021/Week_02_Transit_Cost_Project>

```{r}
library(tidytuesdayR)
library(tidyverse)
library(countrycode)
library(ggbump)
library(ggflags)
```

```{r}
df <- tt_load('2021-01-05')$transit_cost[1:537,]

df$start_year <- as.numeric(df$start_year)
df$end_year <- as.numeric(df$end_year)
df$real_cost <- as.numeric(df$real_cost)
df$country <- countrycode(df$country, origin = 'iso2c', destination = 'genc2c')
```

```{r}

df_country <- df %>%
    group_by(country) %>%
    summarise(total_cost = sum(real_cost), total_length = sum(length)) %>%
    mutate(cost_length = total_cost / total_length) %>%
    slice_max(total_length, n = 20)
```

```{r}

df_rank <- df_country %>%
    mutate(total_length_rank = 1:20) %>%
    arrange(desc(total_cost)) %>%
    mutate(total_cost_rank = 1:20) %>%
    arrange(desc(cost_length)) %>%
    mutate(cost_length_rank = 1:20) %>%
    select(country, ends_with('rank')) %>%
    pivot_longer(cols = ends_with('rank'), names_to = 'metric', values_to = 'rank') %>%
    mutate(metric = factor(metric, levels = c('total_cost_rank', 'cost_length_rank', 'total_length_rank'))) %>%
    mutate(continent = countrycode(country, origin = 'genc2c', destination = 'continent'))
```

```{r fig.width=16, fig.height=12}

ggplot(df_rank, aes(x = metric, y = rank, country = tolower(country), color = continent)) + 
    geom_bump(smooth = 10, size = 1.5, lineend = "round") +
    geom_flag(size = 10) + 
    scale_x_discrete(expand = expansion(mult = c(.05, .05)),
                     labels = c('total_cost_rank' = 'Total cost of lines \n(USD millions)', 'cost_length_rank' = 'Total cost per length of Line \n(USD millions / Km)', 'total_length_rank' = 'Total length of lines \n(Km)')) +
    scale_y_reverse(breaks = seq(1, 20, 1),
                    sec.axis = sec_axis(trans = ~.*1, name = "Second Axis", breaks = seq(1, 20, 1))) + 
    scale_color_brewer(type = 'qual', palette = 'Dark2') +
    labs(title = 'Cost and Length of Transit per Country',
         subtitle = 'Ranking of selected countries in terms of length, cost and cost/length of constructed lines for public transport in the period 1990-2020. \nThe countries shown are the top 20 with the most constructed lines in Km for the 30-year period. \nAsian countries have been investing tons of dollars in transport infrastucture in the past few decades.',
         caption = 'by Andrés Camilo Zúñiga González (@ancazugo) \n Data: Transit Costs Project') +
    theme(
        panel.background = element_rect(fill = '#000000'),
        plot.background = element_rect(fill = '#000000'),
        plot.margin = unit(c(1, 1, 0.5, 1), "cm"),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(color = 'white', size = 20, hjust = 0.5),
        axis.text.y.right = element_text(color = 'white', size = 20, hjust = 0.5),
        axis.text.x = element_text(color = 'white', size = 15),
        plot.title = element_text(face = 'bold', colour = 'white', size = 20),
        plot.subtitle = element_text(colour = 'white', size = 15),
        plot.caption = element_text(colour = 'white', size = 10),
        plot.caption.position = "plot",
        legend.background = element_rect(fill = '#000000'),
        legend.text = element_text(face = 'bold', color = 'white', size = 15),
        legend.position = 'bottom',
        legend.key = element_rect(fill = '#000000'),
        legend.spacing.x = unit(.5, "cm"),
        legend.key.size = unit(2, "cm")
    )
```
