---
title: "Print personalized Street Maps Using R"
output: html_notebook
---

打印个性化的街道地图 refer to: <https://taraskaduk.com/posts/2021-01-18-print-street-maps/>

```{r}
library(sf)
library(osmdata)
library(raster)
library(tidyverse)
library(lwgeom)
```

## 获取数据

利用osmdata来获取Open Street Map数据

```{r}
place <- "Kyiv Ukraine"

highway_sizes <- tibble::tribble(
          ~highway, ~highway_group, ~size,
        "motorway",        "large",   0.5,
   "motorway_link",        "large",   0.3,
         "primary",        "large",   0.5,
    "primary_link",        "large",   0.3,
       "secondary",       "medium",   0.3,
  "secondary_link",       "medium",   0.3,
        "tertiary",       "medium",   0.3,
   "tertiary_link",       "medium",   0.3,
     "residential",        "small",   0.2,
   "living_street",        "small",   0.2,
    "unclassified",        "small",   0.2,
         "service",        "small",   0.2,
         "footway",        "small",   0.2
  )
```

I will be creating a map of Kyiv, Ukraine, but the process should be identical for any other city (except for mapping the bodies of water, which can get tricky)

```{r}
streets_osm <- opq(place) %>%
  add_osm_feature(key = "highway", 
                  value = highway_sizes$highway) %>%
  osmdata_sf()

streets <- streets_osm$osm_lines %>% 
  select(osm_id, name, name.en, highway, maxspeed, oneway, surface) %>% 
  mutate(length = as.numeric(st_length(.))) %>% 
  left_join(highway_sizes, by="highway") %>% 
  filter(highway_group != "small" | length >= quantile(length, probs = 0.25))

railways_osm <- opq(place) %>%
  add_osm_feature(key = "railway", value="rail") %>%
  osmdata_sf()

railways <- railways_osm$osm_lines %>% 
  dplyr::select()
```

For water, a lot depend on how you want to map it. If you only want to show the river as a line, you shouldn't have a problem with key-value pair of "waterway"-"river".

I wanted to map the Dnipro river (the river that runs through Kyiv, as well as through all of Ukraine) as it is on the map. My code below accomplishes that, but it may not work for your body of water. Please do research on key-value pairs for your case then.

```{r}
water_osm <- opq(place) %>%
  add_osm_feature(key = "natural", value = "water") %>%
  osmdata_sf() %>% 
  unname_osmdata_sf()

river_osm <- opq(place) %>%
  add_osm_feature(key = "waterway", value = c("river", "riverbank")) %>%
  osmdata_sf() %>% 
  unname_osmdata_sf()

water <- c(water_osm, river_osm) %>% 
  .$osm_multipolygons %>% 
  select(osm_id, name) %>% 
  mutate(area = st_area(.)) %>% 
  # this filter gets rid of tiny isolated lakes et cetera
  filter(area >= quantile(area, probs = 0.75))
```

## 获得边界

### Go by the city's borders

```{r}
#First, get all admin boundaries associated with your request
all_boundaries <- opq(place) %>%
  add_osm_feature(key = "boundary", 
                  value = c("administrative")) %>%
  osmdata_sf() %>% 
  #This next step is optional, depending on what the call returns
  unname_osmdata_sf() %>% 
  .$osm_multipolygons

ggplot(data = all_boundaries) + 
  geom_sf()
```

```{r}
all_boundaries
```

```{r}
#The previous call returns more than one boundary.
#Inspect the data and choose the object you need.
boundary <- all_boundaries %>% 
  filter(osm_id == 421866) %>% 
  #If you're after geometries only, 
  #you won't need all other 200 some columns
  dplyr::select()
```

### A Circular Map

The algorithm for this approach is to find a center point on the map, then draw a circle around that center point. The inspiration and the code for this approach came from "The Beautiful Hidden Logic of Cities" (Davis 2019).

You will need to know your center's coordinates, the desired radius in meters, as well as your local crs projection (to re-project the coordinates in order to draw a circle with a radius given in meters).

```{r}
crs2 <- 6384 # https://epsg.io/6384

center <- c(long = 30.522,
            lat = 50.451)
center_proj <- tibble(lat = center["lat"], long = center["long"]) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  st_transform(crs = crs2)

dist <-  10000
circle <- tibble(lat = center["lat"], long = center["long"]) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  st_transform(crs = crs2) %>% 
  st_buffer(dist = dist) %>% 
  st_transform(crs = 4326)
```

```{r}
circle
```

### Rectangular Map

If we know a desired center, and know (or guess) a radius, constructing a crop box becomes a matter of drawing a circle around the center, and then finding the tangents on each of four sides.

```{r}
dist <-  10000
circle1 <- tibble(lat = center["lat"], long = center["long"]) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%  
  st_transform(crs = crs2) %>% 
  st_buffer(dist = dist)
bbox1 <- st_bbox(circle1)
```

An extra complication: if we don't want a square, but a rectangular map. I wanted a 4:5 ratio. What to do? Draw another circle, and then use 2 data points from each to form a box.

```{r}
circle2 <- tibble(lat = center["lat"], long = center["long"]) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%  
  st_transform(crs = crs2) %>% 
  st_buffer(dist = dist/0.8) %>% 
  st_transform(crs = 4326)

bbox2 <- st_bbox(circle2)
bbox <- c(bbox1$xmin, bbox2$ymin, bbox1$xmax, bbox2$ymax)
```

If you are going to plot some personal data points on top, this is the time upload your data. Here is a sample I used:

```{r}
points <- tibble::tribble(
             ~Place,      ~Lat,     ~Long, ~Type, ~Person,
            "Abcde", 50.433374, 30.426781,   "A",     "A",
            "Abcde", 50.500973,  30.36365,   "A",     "A",
            "Abcde", 50.456738, 30.622579,   "B",     "A",
            "Abcde", 50.427142, 30.468178,   "C",     "A",
            "Abcde", 50.427348, 30.423886,   "C",     "A",
            "Abcde", 50.425519, 30.436046,   "C",     "A",
            "Abcde", 50.504877, 30.469968,   "D",     "A",
            "Abcde", 50.455176, 30.524202,   "D",     "A",
            "Abcde",  50.42226, 30.425524,   "E",     "A",
            "Abcde", 50.428142, 30.430448,   "E",     "A",
            "Abcde", 50.439139, 30.553577,   "A",     "B",
            "Abcde",  50.40741, 30.401694,   "A",     "B",
            "Abcde", 50.520441,  30.49738,   "A",     "B",
            "Abcde", 50.397799, 30.513417,   "A",     "B",
            "Abcde", 50.447156, 30.431727,   "B",     "B",
            "Abcde", 50.427097, 30.519814,   "B",     "B",
            "Abcde", 50.516017,  30.60801,   "E",     "B",
            "Abcde", 50.509067,  30.46301,   "A",     "C",
            "Abcde", 50.454697, 30.505255,   "C",     "C",
            "Abcde", 50.450976, 30.522568,   "D",     "C",
            "Abcde", 50.420312, 30.528391,   "D",     "C",
            "Abcde", 50.459616, 30.521638,   "D",     "C",
            "Abcde", 50.437684, 30.527156,   "D",     "C",
            "Abcde", 50.457284, 30.438777,   "E",     "C",
            "Abcde", 50.500091, 30.602264,   "E",     "C"
            )
points_sf <- points %>% 
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)
```

\
Another thing that we need to do now is to make sure that all points fit inside the printed map. We're going to find the limits on all sides. Out of curiosity, we also can find the center of such map.

```{r}
top <- max(points$Lat)
bottom <- min(points$Lat)
left <- min(points$Long)
right <- max(points$Long)

center <- c(long = (right - left)/2 + left,
            lat = (top - bottom)/2 + bottom)
```

My approach is slightly different and more complicated though. I have a predetermined center in mind (and I assume you will as well), but I also want to make sure all of my points fit in the map, while center remains center. I need to find the longest of four distances from the center to the top, bottom, left, and right, and use that as my new length. I also will be adding an extra kilometer to the distance to have the furthest point not completely on the edge of the map

```{r}

center <- c(long = 30.5224974,
            lat = 50.4508911)

top1 <- pointDistance(center, c(center["long"], top), lonlat = TRUE)
bottom1 <- pointDistance(center, c(center["long"], bottom), lonlat = TRUE)
right1 <- pointDistance(center, c(right, center["lat"]), lonlat = TRUE)
left1 <- pointDistance(center, c(left, center["lat"]), lonlat = TRUE)

dist <- max(top1,
            bottom1, 
            right1, 
            left1) + 1000 
```

Now that we have a center and a radius, we can draw the box the same way

```{r}
circle1 <- tibble(lat = center["lat"], long = center["long"]) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%  
  st_transform(crs = crs2) %>% 
  st_buffer(dist = dist) %>% 
  st_transform(crs = 4326)

bbox1 <- st_bbox(circle1)

circle2 <- tibble(lat = center["lat"], long = center["long"]) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%  
  st_transform(crs = crs2) %>% 
  st_buffer(dist = dist/0.8) %>% 
  st_transform(crs = 4326)

bbox2 <- st_bbox(circle2)
bbox <- c(bbox1$xmin, bbox2$ymin, bbox1$xmax, bbox2$ymax)
```

### 剪裁地图

With borders determined (one of several ways above), we are ready to crop the map to its desired size. If we've been cutting out a rectangular shape, we can use a st_crop() function to crop the shapefiles. Otherwise, we'll need st_intersection()

If you want to you the city/metro/district/etc borders, the code and the initial output will look something like this:

```{r}
streets_cropped <- streets %>% st_intersection(boundary)
water_cropped <- water %>% st_intersection(boundary)
railways_cropped <- railways %>% st_intersection(boundary)
```

A circle crop will have a very similar code, and will most likely run much faster:

```{r}
streets_cropped <- streets %>% st_intersection(circle)
water_cropped <- water %>% st_intersection(circle)
railways_cropped <- railways %>% st_intersection(circle)
```

A rectangular crop will have a slightly different code, as mentioned before, and will run even faster:

```{r}
water_cropped <- st_crop(water, bbox)
streets_cropped <- st_crop(streets, bbox)
railways_cropped <- st_crop(railways, bbox)
```

## 图像绘制

Now, to the plotting. This is where most of the tweaking is going to happen. Pick colors, opacity, width of the streets, order of layers etc etc. At some point, I wrote a double-layered for loop to iterate over several shades of gray and several versions of street widths, generating 64 different chart combinations. And that's just for 2 parameters alone with values very close to one another! The possibilities here are endless. Have fun customizing the look!

Here is my final version:

```{r fig.width=14, fig.height=18}
blankbg <-theme(axis.line=element_blank(),
                axis.text.x=element_blank(),
                axis.text.y=element_blank(),
                axis.ticks=element_blank(),
                axis.title.x=element_blank(), 
                axis.title.y=element_blank(),
                legend.position = "none",
                plot.background=element_blank(),
                panel.grid.minor=element_blank(),
                panel.background=element_blank(),
                panel.grid.major=element_blank(),
                plot.margin = unit(c(t=2,r=2,b=2,l=2), "cm"),
                plot.caption = element_text(color = "grey20", size = 142, 
                                            hjust = .5, face = "plain", 
                                            family = "Didot"),
                panel.border = element_blank()
)


p <- ggplot() +
  blankbg +
  geom_sf(data = water,
          fill = "steelblue",
          # size = .8,
          lwd = 0,
          alpha = .3) +
  geom_sf(data = railways,
          color = "grey30",
          size = .2,
          linetype="dotdash",
          alpha = .5) +
  geom_sf(data = streets %>% 
            filter(highway_group == "small"),
          size = .1,
          color = "grey40") +
  geom_sf(data = streets %>% 
            filter(highway_group == "medium"),
          size = .3,
          color = "grey35") +
  geom_sf(data = streets %>% 
            filter(highway_group == "large"),
          size = .5,
          color = "grey30") +
  geom_sf(
    data = points_sf,
    aes(#size = Size,
      col = Person),
    alpha = 0.8,
    size = 2
  ) +
  labs(caption = 'Kyiv') +
  coord_sf(xlim = c(bbox[1], bbox[3]),
           ylim = c(bbox[2], bbox[4]),
           expand = FALSE)

p
```
