---
title: "Crop Growth in Kenya by Type and County"
output: html_notebook
---

https://github.com/wjtorres/tidytuesday/blob/master/2021_01_19/2021_01_19_tidy_tuesday.Rmd

```{r}
library(tidyverse)
library(tidytuesdayR)
library(stringr) # to capitalize counties
library(forcats) # for factoring
library(showtext) # for adding google font
```
```{r}
tt <- tt_load("2021-01-19")
```
```{r}
crops <- tt$crops

crops2 <- crops %>%
  
  # remove Kenya
  filter(SubCounty != "KENYA") %>%
  
  #pivot to long format
  pivot_longer(cols = Farming:`Khat (Miraa)`, names_to = "crop" ) %>%
  
  # group by crop and find percentage of each crops grown at the SubCounty
  group_by(crop) %>%
  mutate(percentage = value/sum(value, na.rm = T) * 100) %>%
  ungroup()
# add percentage groups
crops2 <- crops2 %>%
  mutate(percentage_group = case_when(
    percentage > 0 & percentage <= 5 ~ "<5%",
    percentage > 5 & percentage <= 10 ~ "5-10%",
    percentage > 10 & percentage <= 20 ~ "10-20%",
    percentage > 20 ~ ">20%"
    )
  )
# factor reorder the percentage_group
crops2$percentage_group <- fct_reorder(crops2$percentage_group, crops2$percentage, min)
# capitalize first letter only
crops2$SubCounty <- stringr::str_to_title(crops2$SubCounty)
# arrange descending (important for order of appearance)
crops2 <- crops2 %>%
  arrange(desc(SubCounty))
```

```{r}
showtext_auto(enable = TRUE)

# add google fonts
font_add_google(name = "Lustria", family = "one")
font_add_google(name = "Nunito", family = "two")

# save my colors
m_grey <- "#7a7a7a"
d_grey <- "#222222"
# customize theme
my_theme <- theme(
  
  # title
  plot.title = element_text(family = "one", size = 18, hjust = .5, color = d_grey, margin = margin(10, 0, 15, 0)),
  
  # subtitle
  plot.subtitle = element_text(family = "one", size = 11, hjust = .5, color = d_grey, margin = margin(0, 0, 15, 0)),
  
  # caption
  plot.caption = element_text(family = "one", size = 10, hjust = .5, color = m_grey, margin = margin(15, 0, 10, 0)),
  
  
  # legend
  legend.position = "bottom",
  legend.title = element_text(family = "one", size = 10, color = d_grey),
  legend.text = element_text(family = "one", size = 10, color = d_grey),
  legend.background = element_rect(fill = "#d4dddd"),
  legend.key = element_rect(fill = "#d4dddd", color = NA),
  legend.key.width = unit(0.5,"cm"),
  
  # axis
  axis.text = element_text(family = "one", face = "bold", size = 8, color = d_grey),
  axis.ticks = element_blank(),
  
  # panels
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  
  # plot background
  panel.background = element_rect(fill = "#d4dddd", color = "#d4dddd"),
  plot.background = element_rect(fill = "#d4dddd")
)

# create plot
my_plot <- crops2 %>%
  ggplot(aes(x = crop, fct_inorder(SubCounty))) +
  
  # add heat map
  geom_tile(aes(fill = percentage_group), width = 0.8, height = 0.8) +
  
  # add labels
  geom_text(aes(family = "one", label = (round(percentage, 1))), size = 3) +
  
  # scales
  scale_x_discrete(position = "top") +
  scale_fill_manual(values = c("#ece0b6", "#e3d297", "#d7be69", "#cbaa3b"), breaks = c("<5%","5-10%","10-20%",">20%")) +

  # labs
  labs(title = "Crop Growth in Kenya by Type and County",
       subtitle = "For each crop, the numbers represent the percentage of households in the county \n
growing the crop from data available in the 2019 Kenya Population and Housing Census. \n
If data are not available, blanks are used.",
       caption = "Design: Jackie Torres | Data Source: rKenyaCensus by Shelmith Kariuki | #TidyTuesday 2021W4",
       x = NULL, y = NULL, fill = NULL) +
  
  # use my theme
  my_theme
```

```{r fig.width=20, fig.height=30}
my_plot
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
