---
title: "Plastic Pollution"
output: html_notebook
---

refer to <https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-26/readme.md>

## Largest concentration of waste per volunteer

-   <https://twitter.com/AndyOReilly10/status/1355225855736209412?s=09>

-   <https://github.com/mrandyoreilly/TidyTuesday/blob/main/Plastics.Rmd>

```{r}
# load libraries
library(tidyverse)
library(scales)
library(janitor)
library(cowplot)
library(magick)
library(RColorBrewer)
theme_set(theme_light())
```

```{r}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')
```

Cleaning script: shortening country names, getting rid of "EMPTY", pivoting longer, and creating a single "unidentified or other" category:

```{r}
plastics_clean <- plastics %>% 
  filter(country != "EMPTY") %>% 
  mutate(`Unknown` = empty + o) %>% 
  relocate(`Unknown`, .after = pvc) %>% 
  select(-o, -empty) %>% 
  pivot_longer(4:10, names_to = "type", values_to = "count") %>% 
  filter(count != 0) %>% 
  janitor::clean_names() %>% 
  mutate(country = recode(country, "ECUADOR" = "Ecuador", "United Kingdom of Great Britain & Northern Ireland" = "UK",
                          "Taiwan_ Republic of China (ROC)" = "Taiwan", "NIGERIA" = "Nigeria", 
                          "South Africa" = "`S Africa`", "Burkina Faso" = "`B Faso`"))
```

Summarising the data: creating a "per volunteer" variable and trimming out countries that had too few volunteers, or very low tallies per volunteer

```{r}
plastics_summarised <- plastics_clean %>% 
  filter(volunteers >= 30) %>% 
  group_by(country, type) %>% 
  summarise(tallies = sum(count), volunteers = sum(volunteers)) %>% 
  filter(tallies >= 90) %>% 
  mutate(per_volunteer = tallies/volunteers) %>% 
  mutate(country = as.factor(country)) %>% 
  ungroup() %>% 
  group_by(country) %>% 
  mutate(total_per_volunteer = sum(per_volunteer)) %>% 
  filter(total_per_volunteer >= 1.5)
```

First creating a palette that omits the brightest colours. Then making bar plot, then lots and lots and lots of work getting the arrows and annotations right! Finally, background image placed with ggdraw.

```{r fig.width=12, fig.height=7}
my_palette <- brewer.pal(name="GnBu", n=10)[3:10]
 
p  <- plastics_summarised %>% 
  ggplot(aes(x = fct_reorder(country, total_per_volunteer), y = per_volunteer, fill = type)) +
  geom_col(colour = "White") +
  theme_minimal_grid() +
  labs(title = "Largest concentration of waste per volunteer",
       subtitle = "Does Togo have the biggest waste problem, or just the busiest volunteers?",
       y = "Waste items per volunteer",
       x = "",
       fill= "Type of plastic") +
  scale_fill_manual(values = my_palette) +
  theme(panel.grid.major = element_line(linetype = "blank"), 
    panel.grid.minor = element_line(linetype = "blank")) +labs(x = NULL) +
  theme(legend.position = "left") +
  theme(axis.ticks = NULL)  +
  theme(axis.text = NULL) +
  theme(axis.text.x=element_text(angle = -335, hjust = 0.5,  size = 9, colour = "white")) +
  theme(axis.ticks.x = element_blank()) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(size = 20, colour = "White", vjust = -5)) +
  theme(plot.subtitle = element_text(size = 11, colour = "White", vjust = -8.5)) +
  theme(axis.text.y = element_text(colour = "white")) +
  theme(axis.title.y = element_text(colour = "white")) +
  annotate("text", x = 2, y = 12, label = "PET (drink bottles, \nfood containers)", colour = "White") +
  annotate("text", x = 9.5, y = 17, label = "LDPE (Plastic bags, \nsqueeze bottles", colour = "White")+
  annotate("text", x = 10, y = 25, label = "HDPE (Milk containers, \nbottle tops)", colour = "White") +
  annotate("text", x = 6.9, y = 13, label = "PP (Flower pots, \nbeverage cups", colour = "White") +
  annotate("text", x = 4.5, y = 4, label = "PS (Toys, styrofoam)", colour = "White") +
  annotate("text", x = 3.3, y = 20, label = "PVC (Chemical \nbottles, \nplumbing)", colour = "White") +
  annotate("text", x = 5.5, y = 28, label = "Unidentified and other plastics", color = "white") +
  geom_segment(x = 7.5, y = 28, xend = 12, yend = 28, colour = "White", 
                  arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(x = 4, y = 18, xend = 8, yend = 1, colour = "White",
                  arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(x = 10.5, y = 22, xend = 11, yend = 15, colour = "White",
                  arrow = arrow(length = unit(0.3, "cm"))) +
  geom_segment(x = 2, y = 10, xend = 1, yend = 1, colour = "White",
                  arrow = arrow(length = unit(0.3, "cm"))) +
   geom_segment(x = 6, y = 4, xend = 7, yend = 3, colour = "White",
                  arrow = arrow(length = unit(0.3, "cm"))) +
   geom_segment(x = 10, y = 14, xend = 10, yend = 9, colour = "White",
                  arrow = arrow(length = unit(0.3, "cm"))) +
   geom_segment(x = 8, y = 14, xend = 10, yend = 4, colour = "White",
                  arrow = arrow(length = unit(0.3, "cm")))
  
  #theme(plot.background = element_rect(fill = "gray80"))
ggdraw() +
  draw_image("./data/Plastic_bottles_faded.png", halign = 1, valign = 0.2, scale = 2)  +
  draw_plot(p)
```
