---
title: "国省格点预报逐3小时温度预检验评估"
author: 国家气象中心 天气预报技术研发室
date: 2019年5月19日
output:
  html_notebook:
    number_section: true
---

# 简介
  国家级与省级网格预报的24小时时效内、逐3小时间隔温度预报检验评估分析.
  
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
```

# 国家级及省级网格预报绝对误差分析

# 省级相对于国家级预报的负技巧占比
```{r, fig.width=12, fig.height=9}
skill <- read.table('data/20190519_3h_temp_verif/负技巧比例4月.dat', header=FALSE)
names(skill) <- c("name", "03h", "06h", "09h", "12h", "15h", "18h", "21h", "24h")
skill <- gather(skill, "valid.hour", "skill", -name)

# 
g <- ggplot(skill, aes(name, skill))+
  geom_col(fill="#CEA438") +
  facet_grid(valid.hour~.)
g
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
