---
title: "GSODR Data Explore"
author: "Kan Dai"
output: 
  html_notebook:
    toc: true
    toc_depth: 2
---

# Introduction
The GSOD or [Global Surface Summary of the Day (GSOD)](https://data.noaa.gov/dataset/dataset/global-surface-summary-of-the-day-gsod/) data provided by the US National Centers for Environmental Information (NCEI) are a valuable source of weather data with global coverage. GSODR provides four main functions for manipulating this data:

* get_GSOD() - this function queries and transfers files from the NCEI’s FTP server, reformats them and returns a tidy data frame in R

* reformat_GSOD() - this function takes individual station files from the local disk and re-formats them returning a tidy data frame in R

* nearest_stations() - this function returns a vector of station IDs that fall within the given radius (kilometres) of a point given as latitude and longitude

* update_station_list() - this function downloads the latest station list from the NCEI’s FTP server updates the package’s internal database of stations and their metadata.

* get_inventory() - this function downloads the latest station inventory information from the NCEI’s FTP server and returns the header information about the latest version as a message in the console and a tidy data frame of the stations’ inventory for each month that data are reported.

For more inforation see the description of the data provided by NCEI, [http://www7.ncdc.noaa.gov/CDO/GSOD_DESC.txt](http://www7.ncdc.noaa.gov/CDO/GSOD_DESC.txt).

# Using getGSOD()

```{r, warning=FALSE, message=FALSE}
library(GSODR)
library(dplyr)
library(ggplot2)
library(ggalt)
library(ggthemes)
library(maps)
```

  You can download latest station list information ad update internal database
```{r}
update_station_list()
```


## Find Stations in China
GSODR provides lists of weather locations and elevation values. Using dplyr, we can find all the stations in China.

```{r}
load(system.file("extdata", "country_list.rda", package = "GSODR"))
load(system.file("extdata", "isd_history.rda", package = "GSODR"))

station_locations <- left_join(isd_history, country_list, by=c("CTRY" = "FIPS"))

# create data.frame for Australia only
Oz <- subset(station_locations, COUNTRY_NAME == "CHINA")

Oz
```

Seletct stations whose name include "BEI".
```{r}
subset(Oz, grepl("BEI", STN_NAME))
```


## Show the distribution of ground-based weather stations

```{r, warning=FALSE, message=FALSE, fig.height=8, fig.width=12}
world <- map_data("world")
gg <- ggplot()
gg <- gg + geom_map(data=world, map=world, 
                    aes(x=long, y=lat, map_id=region),
                    color="white", fill="#7f7f7f", size=0.05, alpha=0.25)
gg <- gg + geom_point(data=station_locations,
                      aes(x=LON, y=LAT), color="#E69F00", size=0.3, alpha=0.4)
gg <- gg + coord_proj("+proj=wintri")
gg <- gg + labs(title = "The distribution of ground-based weather stations",
                subtitle = paste("Number of stations:", as.character(nrow(station_locations))),
                caption = "Data source: GSOD")
gg <- gg + theme_map()
gg <- gg + theme(strip.background=element_blank())
gg <- gg + theme(legend.position="none")
gg
```


